---
title: "Getting Started with tradeengine"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with tradeengine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction

Welcome to **tradeengine**, the backtesting engine for quantitative trading strategies in R! This vignette will walk you through the basics of creating, testing, and analyzing trading strategies using the TradingVerse framework.

## Philosophy

tradeengine follows the tidyverse philosophy:

- **Pipe-friendly**: Use `|>` or `%>%` to chain operations
- **Consistent API**: Predictable function names and parameters
- **Tidy data**: Everything works with tibbles
- **Flexible**: From simple to complex strategies

## Installation

```{r setup, eval=FALSE}
# Install from GitHub
devtools::install_github("tradingverse/tradeengine")
```

```{r load}
library(tradeengine)
library(dplyr)
```

# The market_tbl: Foundation of TradingVerse

All tradeengine operations start with a `market_tbl` - a standardized tibble for financial time series data.

## Creating Market Data

### From scratch

```{r create_market_tbl}
# Create sample data
prices <- market_tbl(
  symbol = rep("AAPL", 10),
  datetime = seq.POSIXt(as.POSIXct("2024-01-01"), by = "day", length.out = 10),
  open = c(150, 151, 152, 151, 153, 154, 155, 154, 156, 157),
  high = c(152, 153, 154, 153, 155, 156, 157, 156, 158, 159),
  low = c(149, 150, 151, 150, 152, 153, 154, 153, 155, 156),
  close = c(151, 152, 153, 152, 154, 155, 156, 155, 157, 158),
  volume = c(1e6, 1.1e6, 1.2e6, 1.15e6, 1.3e6, 1.25e6, 1.4e6, 1.35e6, 1.5e6, 1.45e6)
)

print(prices)
```

### Using synthetic data

For testing and learning, tradeengine provides a synthetic data generator:

```{r synthetic_data}
# Generate 1 year of realistic data
synthetic <- generate_synthetic_data(
  n_days = 252,
  start_price = 100,
  volatility = 0.20,  # 20% annual volatility
  drift = 0.08,       # 8% annual return
  symbol = "DEMO"
)

head(synthetic, 10)
```

# Building Your First Strategy

## Simple Moving Average Crossover

Let's build a classic SMA crossover strategy:

```{r sma_strategy}
# Add technical indicators
strategy_data <- synthetic |>
  mutate(
    sma_20 = sma(close, 20),
    sma_50 = sma(close, 50)
  )

# Define strategy rules
strategy <- strategy_data |>
  add_strategy(
    name = "SMA_Crossover",
    entry_rules = sma_20 > sma_50,   # Buy when fast crosses above slow
    exit_rules = sma_20 < sma_50,    # Sell when fast crosses below slow
    position_size = 1000,             # $1000 per trade
    commission = 0.001,               # 0.1% commission
    slippage = 0.0005                 # 0.05% slippage
  )

head(strategy, 10)
```

# Running a Backtest

## Vectorized Backtesting

The fastest method for signal-based strategies:

```{r backtest_vectorized}
results <- backtest(
  strategy,
  initial_capital = 10000,
  method = "vectorized"
)

# View summary
summary(results)
```

## Analyzing Results

```{r analyze_results}
# Equity curve
head(results$equity_curve, 10)

# Trade details
if (nrow(results$trades) > 0) {
  head(results$trades, 5)
}

# Key metrics
cat("Final Equity: $", format(results$final_equity, big.mark = ","), "\n")
cat("Total Return: ", round(results$total_return * 100, 2), "%\n")
cat("Number of Trades: ", results$n_trades, "\n")
```

## Visualization

```{r plot_results, eval=FALSE}
# Plot equity curve
plot(results)
```

# Advanced Features

## Position Sizing Methods

tradeengine supports multiple position sizing approaches:

```{r position_sizing}
# Fixed dollar amount
calc_position_size("fixed", capital = 10000, price = 100, risk_per_trade = 0.02)

# Percentage of equity
calc_position_size("percent_equity", capital = 10000, price = 100, risk_per_trade = 0.05)

# Volatility targeting
calc_position_size(
  "volatility_target", 
  capital = 10000, 
  price = 100,
  volatility = 0.30,        # Asset volatility
  target_volatility = 0.10  # Target portfolio volatility
)
```

## Portfolio Management

Create and backtest multi-asset portfolios:

```{r portfolio, eval=FALSE}
# Create a 60/40 stock/bond portfolio
portfolio <- create_portfolio(
  symbols = c("SPY", "TLT"),
  weights = c(0.6, 0.4),
  rebalance_frequency = "monthly",
  initial_capital = 100000
)

# Backtest portfolio
portfolio_results <- backtest(portfolio)
```

## Complex Strategy Rules

You can create sophisticated multi-condition strategies:

```{r complex_strategy, eval=FALSE}
# Multi-indicator strategy
complex_strategy <- synthetic |>
  mutate(
    sma_20 = sma(close, 20),
    sma_50 = sma(close, 50),
    returns = calc_returns_simple(close)
  ) |>
  add_strategy(
    name = "Multi_Indicator",
    entry_rules = (sma_20 > sma_50) & (returns > 0) & (volume > lag(volume)),
    exit_rules = (sma_20 < sma_50) | (returns < -0.05),  # Exit or stop loss
    position_size = 2000
  )
```

# Best Practices

## 1. Always Include Transaction Costs

Real trading has costs. Always model them:

```{r transaction_costs, eval=FALSE}
strategy <- data |>
  add_strategy(
    entry_rules = ...,
    exit_rules = ...,
    commission = 0.001,  # 0.1% per trade
    slippage = 0.0005    # 0.05% slippage
  )
```

## 2. Use Realistic Position Sizing

Don't risk your entire capital on one trade:

```{r realistic_sizing, eval=FALSE}
# Risk 2% of capital per trade
position_size_func <- function(equity, price) {
  risk_amount <- equity * 0.02
  shares <- risk_amount / price
  return(floor(shares))
}
```

## 3. Validate Your Data

Always check your market data:

```{r validate, eval=FALSE}
# Check for issues
validate_market_tbl(prices)

# Look for gaps
summary(prices)
```

## 4. Start Simple, Then Iterate

Begin with simple strategies and add complexity gradually:

1. Single indicator (SMA)
2. Two indicators (SMA crossover)
3. Multiple conditions
4. Position sizing rules
5. Portfolio allocation

# Next Steps

- **Strategy Design**: Learn advanced strategy patterns
- **Transaction Costs**: Understand slippage and market impact
- **Portfolio Management**: Multi-asset backtesting
- **Walk-Forward Analysis**: Out-of-sample testing

# Integration with TradingVerse

tradeengine is part of a larger ecosystem:

```{r ecosystem, eval=FALSE}
library(tradeio)        # Data acquisition
library(tradefeatures)  # Technical indicators  
library(tradeviz)       # Visualization
library(trademetrics)   # Performance analytics

# Complete workflow
results <- fetch_prices("AAPL", from = "2023-01-01") |>  # tradeio
  add_rsi(14) |>                                         # tradefeatures
  add_strategy(entry = rsi < 30, exit = rsi > 70) |>    # tradeengine
  backtest(initial_capital = 10000) |>                   # tradeengine
  calc_sharpe() |>                                       # trademetrics
  plot_equity_curve()                                    # tradeviz
```

# Disclaimer

⚠️ **Important**: This software is for educational and research purposes only. Past performance does not guarantee future results. Always do your own due diligence before trading with real money.

# Resources

- [GitHub Repository](https://github.com/tradingverse/tradeengine)
- [API Documentation](https://tradingverse.github.io/tradeengine)
- [TradingVerse Homepage](https://tradingverse.org)
- [Community Forum](https://discord.gg/tradingverse)

---

*Happy trading! 📈*
